---
title: "Class 11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database

The [PDB](https://www.rcsb.org) is the main repository for biomolecular structure data. 

Here we examine the conents of the PDB:

> Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. From
the website what proportion of structures are protein? Aim to have a rendered GitHub document
with working code that yields your answers.

```{r}
PDB_data_exp <- read.csv("PDB Data Export Summary.csv", row.names=1)
View(PDB_data_exp)
```

How many are X-Ray?

```{r}
(PDB_data_exp$Total/sum(PDB_data_exp$Total))*100
```

What percent are Protein?

```{r}
sum(PDB_data_exp$Proteins)/sum(PDB_data_exp$Total) * 100
```


We could also try the datapasta package and copy from website and Addins > DATAPASTA > paste data.frame:
```{r}
library(datapasta)

tmp <- data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(126880, 11062, 2277, 256, 129, 140604),
         Nucleic.Acids = c(2012, 1279, 31, 4, 5, 3331),
    Protein.NA.Complex = c(6547, 259, 800, 6, 2, 7614),
                 Other = c(8, 8, 0, 13, 1, 30),
                 Total = c(135447, 12608, 3108, 279, 137, 151579)
)

View(tmp)
```


> Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

There are 1157 as of 2019-05-07: http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=9DD6CD99

## 3.1 Getting Started with Bio3D
```{r}
library(bio3d)

pdb <- read.pdb("1hsg")

pdb
```

> Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein
residues?

There are 198 amino acid residues in 1HSG. The two nonprotein residues are water and MAP kinase 1, the ligand. 

```{r}
attributes(pdb)

head(pdb$atom)

```

> Q7. What type of R object is pdb$atom? HINT: You can always use the str() function to get a
useful summery of any R object.

pdb$atom is a data frame with numerical, character, and integer variables. 

```{r}
str(pdb$atom)
```

```{r}
# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

Atom selection is done via the function **atom.select()**
```{r}
inds <- atom.select(pdb, "ligand")
```
```{r}
pdb$atom[inds$atom, ]
```

```{r}
prot.pdb <- atom.select(pdb, "protein", value = TRUE)
write.pdb(prot.pdb, file = "1hsg_protein.pdb")
```

```{r}
lig.pdb <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(lig.pdb, file = "1hsg_ligand.pdb")
```

## 3.1 Getting Started with Bio3D

```{r}
library(bio3d)
```

Use help(bio3d) for help.
Use lbio3d() to list bio3d functions, and use help("<function name>") for more information about a particular function.

## 3.2 Bio3D Functions and Common Uses

Use example(<function>) to show examples of a Bio3D function usage:
```{r}
library(bio3d)
example(vmd)
```

```{r}
example(plot.bio3d)
```

## 3.3 Working with individual PDB files

