---
title: "Class 12: Structural Bioinformatics II"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.1 Obtaining and inspecting our input structure
>Load the Bio3D package and use the get.pdb() function to download the "1hsg" PDB entry into your RStudio Project directory:

```{r}
library(bio3d)

pdbid <- get.pdb("1HSG")
hiv <- read.pdb(pdbid)

# summary of our read-in
hiv
```

# Q1
>What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all residue values in this structure?

The two non-protein resids are HOH (water) and MK1 (indinavir).


==========

```{r}
# read in protein only
prot <- atom.select(hiv, "protein", value = TRUE)
write.pdb(prot, file = "1hsg_protein.pdb")

# read in ligand only
lig <- atom.select(hiv, "ligand", value = TRUE)
write.pdb(lig, file = "1hsg_ligand.pdb")

```

>Use the trim.pdb() function to make protein-only and ligand only objects called prot and lig that you can then write out to new PDB format files in your RStudio project directory.

```{r}
# trim.pdb(prot, file = "1hsg_protein.pdb")
# trim.pdb(lig, file = "1hsg_ligand.pdb")
```

## 1.3 Using AutoDockTools to setup protein docking input
>Docking algorithms require each atom to have a charge and an atom type that describes its properties. However, typical PDB structures don’t contain this information. We therefore have to ‘prep’ the protein and ligand files to include these values along with their atomic coordinates. All this will be done in a tool called AutoDock Tools (ADT for short).

# Q2
>Can you locate the binding site visually? Note that crystal structures normally lack hydrogen atoms, why?

The binding site appears to be a vacant space inside the structure, which lacks hydrogen atoms because they're smaller than the x-ray crystallography resolution.

# Q3
>Q3: Look at the charges. Does it make sense (e.g. based on your knowledge of the physiochemical properties of amino acids)?


## Process the **all.pdbtq** file to a PDB format that can be read by VMD (use the Bio3D package)
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```


## Normal Mode analysis for flexibility prediction
```{r}
pdb <- read.pdb("1HEL")
modes <- nma(pdb)
plot(modes, sse=pdb)

# Visualize NMA results
mktrj(modes, mode=7, file="nma_7.pdb")
```

