---
title: "BIBC103 (§A01): Fibroblast Growth Factor"
output: github_document
author: "Ezequiel Quevedo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Normal Mode analysis for flexibility prediction of ternary FGF2-FGFR1-Heparin complex
>DOI: 10.2210/pdb1FQ9/pdb
Classification: GROWTH FACTOR/GROWTH FACTOR RECEPTOR
Organism(s): Homo sapiens
Expression System: Escherichia coli
Mutation(s): 3 

>Deposited: 2000-09-04 Released: 2000-09-27 
Deposition Author(s): Schlessinger, J., Plotnikov, A.N., Ibrahimi, O.A., Eliseenkova, A.V., Yeh, B.K., Yayon, A., Linhardt, R.J., Mohammadi, M.

```{r}
library(bio3d)

pdb <- read.pdb("1FQ9")
modes <- nma(pdb)
plot(modes, sse = pdb)

# Visualize NMA results
mktrj(modes, mode = 7, file = "1fq9_nma_7.pdb", pdb=pdb)
```

## Normal Mode analysis for flexibility prediction of FGF2-FGFR1 complex
>DOI: 10.2210/pdb1CVS/pdb
Classification: GROWTH FACTOR/GROWTH FACTOR RECEPTOR
Organism(s): Homo sapiens
Expression System: Escherichia coli
Mutation(s): 3 

>Deposited: 1999-08-24 Released: 2000-01-28 
Deposition Author(s): Plotnikov, A.N., Schlessinger, J., Hubbard, S.R., Mohammadi, M.
Experimental Data Snapshot

>Method: X-RAY DIFFRACTION
Resolution: 2.8 Å
R-Value Free: 0.281 
R-Value Work: 0.240 

```{r}
library(bio3d)

pdb <- read.pdb("1CVS")
modes <- nma(pdb)
plot(modes, sse = pdb)

# Visualize NMA results
mktrj(modes, mode = 7, file = "1cvs_nma_7.pdb", pdb = pdb)
```




## Normal Mode analysis for flexibility prediction FGF2-Disaccharide (S3I2) complex
>DOI: 10.2210/pdb4OEE/pdb
Classification: PROTEIN BINDING
Organism(s): Homo sapiens
Expression System: Escherichia coli BL21
Mutation(s): 2 

>Deposited: 2014-01-13 Released: 2014-07-09 
Deposition Author(s): Li, Y.C., Hsiao, C.D.
Experimental Data Snapshot

>Method: X-RAY DIFFRACTION
Resolution: 1.5 Å
R-Value Free: 0.185 
R-Value Work: 0.163 

```{r}
library(bio3d)

pdb <- read.pdb("4OEE")
modes <- nma(pdb)
plot(modes, sse = pdb)

# Visualize NMA results
mktrj(modes, mode = 7, file = "4OEE_nma_7.pdb", pdb = pdb)
```